### 2021-06-14 abstractESSLLI21/abstractESSLLI1.Rmd
revised abstract

\footnotesize
\begin{align}
\mathsf{S}(t_i, A_j)  = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,
\mathsf{MAC}(T,A)  = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j) \tag{MAC}
\end{align}
\normalsize 
\noindent That is, for each protected word  and each attribute class, they first take the mean for this protected word and all attributes in a given attribute class, and then take the mean of thus obtained means for all the protected words.

Such methods are statistically problematic. Two issues, specific to @Manzini2019blackToCriminal, are that (i) no distinction is made based on whether given a protected word and a  class of attributes is stereotypically associated with this  protected word or a different protected word, and (ii) no re-scaling by standard deviation (typical in effect size measures such as  Cohen's $d$) is used, so MAC is not a good measure of effect size. A more general problem, is that all the authors ignore the step of comparing their results with control groups, especially control groups of stereotype-neutral human attributes.  That such a move is important is suggested for instance by Figure 1 prepared using the original word list for religion-related protected words extended with control attributes, where such human attributes are also closer to protected classes than neutral non-human words.

Another, also serious problem, is that  all the authors calculate means of means and  run statistical tests on sets of means. Unfortunately, by pre-averaging the data they throw away information about sample sizes, and  they remove variation, and so pre-averaging  tends to manufacture false confidence.

All such approaches come up with rather short lists of protected words and rather short lists of stereotypical attributes. Clearly, these are not complete list. So let's treat them as  samples from richer pools of stereotypical predicates and let's take the uncertainty and variation involved seriously. To illustrate, let's employ the formulas used by @Caliskan2017semanticsBiases in a simple simulation.

>>>>

\footnotesize
\begin{align}
\mathsf{S}(t_i, A_j)  = \frac{1}{\vert A_j\vert}\sum_{a\in A_j}\mathsf{cos}(t,a) & \,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,
\mathsf{MAC}(T,A)  = \frac{1}{\vert T \vert \,\vert A\vert}\sum_{t_i \in T }\sum_{A_j \in A} S(t_i,A_j) \tag{MAC}
\end{align}
\normalsize 
\noindent That is, for each protected word  and each attribute class, they first take the mean for this protected word and all attributes in a given attribute class, and then they take the mean of thus obtained means for all the protected words.

Such methods are statistically problematic. Two issues, specific to @Manzini2019blackToCriminal, are that (i) no distinction is made based on whether  a  class of attributes is stereotypically associated with a given   protected word or with a different protected word, and (ii) no re-scaling by standard deviation (typical in effect size measures such as  Cohen's $d$) is used, so MAC is not a good measure of effect size. A more general problem, is that all the authors ignore the step of comparing their results with control groups, especially control groups of stereotype-neutral human attributes.  That such a move is important is suggested for instance by Figure 1 prepared using the original word list for religion-related protected words extended with control attributes, where such human attributes are also closer to protected classes than neutral non-human words.

Another  serious problem is that  all the authors calculate means of means and  run statistical tests on sets of means. Unfortunately, by pre-averaging the data they throw away information about sample sizes, and  they remove variation, and so pre-averaging  tends to manufacture false confidence.

To illustrate, let's employ the formulas used by @Caliskan2017semanticsBiases in a simple simulation. 
All such approaches come up with rather short lists of protected words and rather short lists of stereotypical attributes. Clearly, these are not complete list. So let's treat them as  samples from richer pools of stereotypical predicates and let's take the uncertainty and variation involved seriously.

### 2021-06-14 abstractESSLLI21/abstractESSLLI1.tex
revised abstract

\normalsize  \noindent That is, for each protected word and each
attribute class, they first take the mean for this protected word and
all attributes in a given attribute class, and then take the mean of
thus obtained means for all the protected words.

Such methods are statistically problematic. Two issues, specific to
{[}6{]}, are that (i) no distinction is made based on whether given a
protected word and a class of attributes is stereotypically associated
with this protected word or a different protected word, and (ii) no
re-scaling by standard deviation (typical in effect size measures such
as Cohen's \(d\)) is used, so MAC is not a good measure of effect size.
A more general problem, is that all the authors ignore the step of
comparing their results with control groups, especially control groups
of stereotype-neutral human attributes. That such a move is important is
suggested for instance by Figure 1 prepared using the original word list
for religion-related protected words extended with control attributes,
where such human attributes are also closer to protected classes than
neutral non-human words.

Another, also serious problem, is that all the authors calculate means
of means and run statistical tests on sets of means. Unfortunately, by
pre-averaging the data they throw away information about sample sizes,
and they remove variation, and so pre-averaging tends to manufacture
false confidence.

All such approaches come up with rather short lists of protected words
and rather short lists of stereotypical attributes. Clearly, these are
not complete list. So let's treat them as samples from richer pools of
stereotypical predicates and let's take the uncertainty and variation
involved seriously. To illustrate, let's employ the formulas used by
{[}2{]} in a simple simulation.

>>>>

\normalsize  \noindent That is, for each protected word and each
attribute class, they first take the mean for this protected word and
all attributes in a given attribute class, and then they take the mean
of thus obtained means for all the protected words.

Such methods are statistically problematic. Two issues, specific to
{[}6{]}, are that (i) no distinction is made based on whether a class of
attributes is stereotypically associated with a given protected word or
with a different protected word, and (ii) no re-scaling by standard
deviation (typical in effect size measures such as Cohen's \(d\)) is
used, so MAC is not a good measure of effect size. A more general
problem, is that all the authors ignore the step of comparing their
results with control groups, especially control groups of
stereotype-neutral human attributes. That such a move is important is
suggested for instance by Figure 1 prepared using the original word list
for religion-related protected words extended with control attributes,
where such human attributes are also closer to protected classes than
neutral non-human words.

Another serious problem is that all the authors calculate means of means
and run statistical tests on sets of means. Unfortunately, by
pre-averaging the data they throw away information about sample sizes,
and they remove variation, and so pre-averaging tends to manufacture
false confidence.

To illustrate, let's employ the formulas used by {[}2{]} in a simple
simulation. All such approaches come up with rather short lists of
protected words and rather short lists of stereotypical attributes.
Clearly, these are not complete list. So let's treat them as samples
from richer pools of stereotypical predicates and let's take the
uncertainty and variation involved seriously.
